---
title: "Arnhold BWMPA Project - Data wrangling"
author: "Gavin McDonald (emLab)"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document: 
    number_sections: yes
    toc: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
params:
     download_data: FALSE
     aggregate_data: FALSE
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = ifelse(Sys.info()["sysname"]=="Windows","G:/Shared\ drives/emLab/Projects/current-projects/arnhold-bwmpa-redistribution/project-materials/markdown-reports",
                         "/Volumes/GoogleDrive/Shared\ drives/emLab/Projects/current-projects/arnhold-bwmpa-redistribution/project-materials/markdown-reports")) })
---

```{r echo = FALSE}
# This chunk sets up default settings for all chunks below
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,fig.width = 7.5,fig.height = 5,dev = 'png',dpi=300)
```

```{r include=FALSE}
# Load all necessary packages
library(tidyverse)
library(sf)
library(stars)
library(rerddap)
library(rnaturalearth)
library(glue)
library(scico)
library(lubridate)
library(furrr)
source(here::here("r/_functions_data_wrangling_erddap.R"))
# Set the data directory. This specifies the folder in Team Drive where the data can be found. This code automatically detects the system type and assigns the correct directory
# data_directory <- ifelse(Sys.info()["sysname"]=="Windows","G:/Shared\ drives/emLab/Projects/current-projects/arnhold-bwmpa-redistribution/data",
#                          "/Volumes/GoogleDrive/Shared\ drives/emLab/Projects/current-projects/arnhold-bwmpa-redistribution/data")
# 
# figure_directory <- ifelse(Sys.info()["sysname"]=="Windows","G:/Shared\ drives/emLab/Projects/current-projects/arnhold-bwmpa-redistribution/project-materials/figures",
#                            "/Volumes/GoogleDrive/Shared\ drives/emLab/Projects/current-projects/arnhold-bwmpa-redistribution/project-materials/figures")
data_directory <- "/Users/gmcdonald/Library/CloudStorage/GoogleDrive-gmcdonald@ucsb.edu/Shared\ drives/emlab/Projects/current-projects/arnhold-bwmpa-redistribution/data"
# Set ggplot theme for all plots
theme_set(theme_minimal() +
            theme(axis.title.y = element_text(angle = 0,vjust=0.5),
                  strip.background = element_blank(),
                  strip.text.y = element_text(angle=0),
                  strip.text.y.right = element_text(angle=0),
                  strip.text.y.left = element_text(angle=0),
                  panel.grid = element_blank(),
                  panel.background = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.grid.major = element_blank()))

# Set mapping projection
map_projection <- "+proj=eqearth +datum=WGS84 +wktext"
# Create global land sf object for mapping
world_plotting <- ne_countries(scale = "small", returnclass = "sf")  %>%
  dplyr::select(geometry)

# Get high-res ocean data
ocean <- ne_download(scale = 50, type = 'ocean', category = 'physical',returnclass = "sf") %>%
  dplyr::select(geometry)
```

# Defining our global grid

```{r}
pixel_size <- 1

# Create polygon rectangle of globe
# This will serve as basis of grid
# global_polygon_sf <- 
#   tibble(lon = c(-180,-180,180,180,-180), 
#        lat = c(-90,90,90,-90,-90))%>% 
#   as.matrix() %>%
#   list(.) %>%
#   st_polygon() %>%
#   st_sfc(crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
#   st_as_sf()+

# global_polygon_sf_small <- tibble(lon = c(-5,-5,5,5,-5),
#                                   lat = c(-5,5,5,-5,-5))%>%
#   as.matrix() %>%
#   list(.) %>%
#   st_polygon() %>%
#   st_sfc(crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs") %>%
#   st_as_sf()

# Start with global ocean
starting_shape <- ocean
#starting_shape <- global_polygon_sf_small
```

```{r eval = params$download_data}
# Make a grid using desired pixel size
data_grid <- starting_shape %>%
  make_grid_custom(pixel_size = pixel_size)

data_grid %>%
  # Add geometry wkt column for saving this as a csv
  mutate(geometry_wkt = st_as_text(geometry)) %>% 
  # Make geometry point, grab lon/lat in lower left-hand corner
  # We'll use this later for joining to GFW data
  st_cast("POINT") %>%
  dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])%>%
  st_set_geometry(NULL)  %>% 
  group_by(pixel_id,geometry_wkt) %>% 
  summarize(lon = min(lon),
            lat = min(lat))%>%
  ungroup()  %>%
  as_tibble() %>%
  write_csv(here::here("data/model_features/global_grid.csv"))
```

We define a global grid using pixels that are `r pixel_size` degree latitude by `r pixel_size` degree longitude. We will then aggregate all of our model feature data to these `r pixel_size`x`r pixel_size` degree pixels. We include only pixels that overlap with the ocean. Each pixel is assigned a static `pixel_id`.


```{r}
data_grid <- data.table::fread(here::here("data/model_features/global_grid.csv")) %>%
  st_as_sf(wkt = "geometry_wkt",
           crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
```

\pagebreak

# Environmental model features

## SST and SST anomaly data 

We use 0.25 degree data from [SST, Daily Optimum Interpolation (OI), AVHRR Only, Version 2.1, Final, Global, 0.25°, 1981-present, Lon+/-180](https://coastwatch.pfeg.noaa.gov/erddap/info/ncdcOisst21Agg_LonPM180/index.html). These data download ~25x as fast and are much faster to spatially aggregate.

```{r eval = params$download_data}
# Download all global SST and SST anomaly data
download_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                             variables = c("sst","anom"),
                             temporal_resolution = "day",
                             date_start = "2012-01-01",
                             date_end = "2021-12-31",
                             data_grid = data_grid,
                             run_parallel = TRUE)
```


```{r eval = params$aggregate_data}
# Are there missing dates from our downloads?
missing_dates <- determine_missing_dates(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_resolution = "day",
                                         date_start = "2012-01-01",
                                         date_end = "2021-12-31")
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data
spatially_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                        spatial_aggregation = data_grid,
                                        run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data
temporally_aggregate_errdap_data_wrapper(dataset_name = "ncdcOisst21Agg_LonPM180",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE)%>%
    # Write aggregated data to clean data folder
    data.table::fwrite(here::here("data/model_features/errdap_sst.csv"))
```

\pagebreak

# Chlorophyll

We get monthly chlorophyll data from [Chlorophyll-a, Aqua MODIS, NPP, L3SMI, Global, 4km, Science Quality, 2003- present (Monthly Composite)](https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chlamday.html). Units are mg m-3 .


```{r eval = params$download_data}
# Download all global  chlorp data
download_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                             variables = c("chlorophyll"),
                             temporal_resolution = "month",
                             date_start = "2012-01-01",
                             date_end = "2021-12-31",
                             data_grid = data_grid,
                             run_parallel = TRUE)
```

```{r eval = params$aggregate_data}
# Are there missing dates from our downloads?
missing_dates <- determine_missing_dates(dataset_name = "erdMH1chlamday",
                                         temporal_resolution = "month",
                                         date_start = "2012-01-01",
                                         date_end = "2021-12-31")
```

```{r eval = params$aggregate_data}
knitr::kable(missing_dates$missing_dates_summary)
```

```{r eval = params$aggregate_data}
# Spatially aggregate chl data
spatially_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                        spatial_aggregation = data_grid,
                                        run_parallel = FALSE)
```

```{r eval = params$aggregate_data}
# Spatially aggregate SST and SST anomaly data
temporally_aggregate_errdap_data_wrapper(dataset_name = "erdMH1chlamday",
                                         temporal_aggregation = "year",
                                         run_parallel = TRUE)%>%
    # Write aggregated data to clean data folder
    data.table::fwrite(here::here("data/model_features/errdap_chl.csv"))
```